---
title: "MLE_NJ"
author: "Zhe Zheng"
date: "3/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# library
library(deSolve)
library(RColorBrewer)
library(reshape2)
```

```{r}
#T is N time points; N_ages is number of age groups
yinit <- readRDS('C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/NJ/yinit.rds') #intial population
birth <-  readRDS('C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/NJ/BrNJ.rds')
contactUSAinfant <- readRDS( 'C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/contactmatrix.rds') # contact matrix
nj_agedist <- readRDS('C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/NJ/nj_agedist.rds')
rsv_nj <- readRDS('C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/NJ/rsv_nj.rds')
```

#View contact matrix
```{r}
n.cols=100
nice.cols <-  colorRampPalette(brewer.pal(9, "YlOrRd"))(n.cols)
heatmap(contactUSAinfant/sum(diag(contactUSAinfant)), Rowv=NA, Colv=NA, scale='none', col=nice.cols)
```

Set names of Age groups 
```{r}
N_ages <- nrow(yinit) 
agenames <- c("<2m","2-3m","4-5m","6-7m","8-9m","10-11m","1Y","2-4Y","5-9Y","10-19Y","20-39Y","40-59Y","60Y+")
al <- N_ages
```

## Population size by time
```{r}
p <- sum(yinit)  # Total population at each time, a vector of length T
```

##Settings parameters
Most of these are described in Table 2 of Pitzer et al, PLOS Pathogens: https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1004591

For birth rate:
The weekly per capita birth rate (B_w) is equal to log(1+B)/52.18 because the data on the birth rate is annual and the differential equation model inherently assumes that growth is occurring exponentially. So if the annual birth rate is equal to 12 per 1000 per year (B=0.012), for example, then we are assuming that:

$N1=N0exp(B_w*52.10)$     where N1 is the population after 1 year and N0 is the baseline population, and $N1=N0*(1+B)$.

$1.012=1*exp(B_w*52.18)$

$B_w=log(1.012)/52.18$.

```{r setparms}
# initialize populationinput matrix as a vector
rownames(yinit) <- agenames
yinit.vector <- as.vector(yinit)

name.array <- array(NA, dim=dim(yinit))
for(i in 1:dim(name.array)[1]){
  for(j in 1:dim(name.array)[2]){
      name.array[i,j] <- paste(dimnames(yinit)[[1]][i],dimnames(yinit)[[2]][j]  )
    }
  }

name.vector <- as.vector(name.array)
names(yinit.vector) <- name.vector
#########################################
#Relative infectiousness for 2nd and subsequent infections
rho1 = 0.75
rho2 = 0.51
#########################################

#########################################
# duration of infectiousness (months)
#Duration in days
dur.days1 <- 10 #days
dur.days2 <- 7 #days
dur.days3 <- 5 #days
###########################################  

###########################################
WidthAgeClassMonth = c(rep(2,times=6), 12,12*3,  60, 120, 240, 240, 240 )  #Aging rate=1/width age class (months) Vector of long N_age
############################################

############################################

#um : calibrated this parameter so we can reproduce the population growth
um= -0.000025
#############################################

#############################################
#Birth rate (births/person/YEAR
#Matrix: T rows, N_ages columns; columns 2:N_ages all 0s
PerCapitaBirthsYear=birth
#############################################

#############################################
#Relaive risk of infection following 1st, 2nd, 3rd+ infections
sigma1=0.76
sigma2=0.6
sigma3=0.4
#############################################

#############################################
#Relaive risk of hospitalizations following 1st, 2nd infections
hosp1= c(0.082251950, 0.047607003, 0.026990367, 0.016398471, 0.012376537, 0.012007695, 0.009609524,0.006999893, 0.001000000, 0.001000000, 0.001000000, 0.001000000, 0.001000000)
hosp2= 0.4*hosp1
#############################################

#############################################
#rate of recovery following 1st, 2nd, 3rd+ infections
#############################################
length.step=30.44 #days per month

gamma1= 1/(dur.days1/length.step)  #converts 1/days to 1/lenth.step
  gamma2= 1/(dur.days2/length.step)  
  gamma3= 1/(dur.days3/length.step)  

```

```{r}
 start_time = 1 # start date (years)
 tmax = (2014-1980)*12
# end_time = 25 # end date (years)
nj_times <- seq(start_time, tmax, by = 1) # gives a sequence from start to end
											 # in increments of 1
```

## Read in transmission dynamic model
```{r}
source("C:/Users/70914/Box/aim3/RSVtransmissionmodel/Model_source/Novac_dynamics.R")
```

```{r saveparms}
parmstest<-list(PerCapitaBirthsYear=PerCapitaBirthsYear,
            DurationMatImmunityDays=112,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            b1=0.15,
            phi=3.35,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit,
            baseline.txn.rate = 10,
            q=1,
            c2=contactUSAinfant,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month'
            )
```

## Run the model 
Note here we are just simulating with set parameters, not fitting to data
```{r}
results_test <- ode(y=yinit.vector, t=nj_times,  
               func=Novac_dynamics, 
            parms=parmstest)

``` 

```{r}
plot(rowSums(results_test[,-1]),type="l") # make sure um reproduce population growth
rowSums(results_test[,-1])[12*25]# should equals to about 19.33 million
```

## Extract the relevant results to see if the epidemic reach equilibrium
```{r}
#used a 25-YEARS burn in period
burnN <- 25*12
results.burned <- results_test[-c(1:burnN),]

##Any infected person
infected.cols <- results.burned[,c(grep('I1', colnames(results.burned)),grep('I2', colnames(results.burned)),grep('I3', colnames(results.burned)),grep('I4', colnames(results.burned)  )) ]

infected.all <- apply(infected.cols,1, sum)

infected.cols.m <- melt(infected.cols)
infected.cols.m$agegrp <- sub(" .*", "",infected.cols.m$Var2 )
infected.cols.c <- dcast(infected.cols.m, Var1~agegrp, fun.aggregate = sum)

plot(infected.all, type='l')

```

## fit the model to NJ inpatient data in order to estimate inital value of parameters
```{r}
parmset<-list(PerCapitaBirthsYear=PerCapitaBirthsYear,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit,
            q=1,
            c2=contactUSAinfant,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month'
            )
```


## fit transmission models to get inputs for MCMC process
```{r}
fitmodel <-  function(parameters,dat) {   # takes the parameter values and dataset as inputs
  protrans <- parameters[1] # estimate parameter related to R0 (baseline transmission rate)
  amp <- parameters[2] # estimate parameter related to seasonal amplitude
  trans <- parameters[3] # estimate parameter related to seasonal peak timing
  DMD <- parameters[4] # estimate parameter related to the duration of maternal immunity
  report_ratio <- exp(parameters[5])
  b1 <- exp(amp) #ensure positive 
  baseline.txn.rate <- exp(protrans) #ensure positive
  phi <-  (2*pi*(exp(trans))) / (1+exp(trans)) # transform to its scale in the model
  durx <- exp(DMD) #ensure positive

   # Simulate the model with initial conditions and timesteps defined above, and parameter values from function call
  results <- ode(y=yinit.vector, t=nj_times,  
               func=Novac_dynamics, 
            parms=c(parmset,baseline.txn.rate=baseline.txn.rate,b1=b1,phi=phi,DurationMatImmunityDays=durx))
  t0 <- 108
  burnN <- 25*12
  St <- results[-c(1:burnN),-1]
   
  I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
     
   hosp1 <- hosp1
   hosp2 <- hosp2
   hosp3 <- rep(0,13)
   # older adults have higher risk of requiring inpatiant care 
   # https://academic.oup.com/jid/article/206/1/56/834700?login=true
   
    beta <-  baseline.txn.rate/(parmset$dur.days1/30.44)/(sum(parmset$yinit.matrix)^(1-parmset$q))*parmset$c2# q depends on transmission type (whether depends on population density or not); c2 is the contact matrix

    lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}
    
      H1=matrix(0,nrow=t0,ncol=al)#Number of hospitalizations by age
      for (i in 1:al){
     H1[,i]=hosp1[i]*S0[,i]*lambda1[,i]+hosp2[i]*sigma1*S1[,i]*lambda1[,i]+hosp3[i]*sigma2*S2[,i]*lambda1[,i]+hosp3[i]*sigma3*S3[,i]*lambda1[,i]}
  H <- report_ratio*rowSums(H1)
    age_dist <- colSums(H1)
 
   LLall <- sum(dpois(x = dat, lambda =H, log = TRUE)) # fit to timeseries
   LLmulti <- dmultinom(x= age_dist,prob = nj_agedist,log = T) # fit to age distribution
   ratio_prior <- dbeta(x=report_ratio,1,1,log=T)
   #total LL
   LL <- LLall+LLmulti+ratio_prior
   
   return(LL)
}

fitLL <- optim(par = c(2.082861e+00, -1.875327e+00, -1.284655e+00,  3.684966e+00, -1.396152e-05),           # 
               #starting values for beta and gamma - you should get the same result no matter which values you choose here
      fn = fitmodel,        # the distance function to optimise
      dat = rsv_nj$H,         # the dataset to fit to ("dat" argument is passed to the function specified in fn)
      control = list(fnscale=-1)) # the log likelihood is negative; here we maximize the log likelihood
```


### Refit model with estimated parameters
```{r,eval=F}
  baseline.txn.rate=exp(fitLL$par[1])
  b1=exp(fitLL$par[2])
  phi=(2*pi*(exp(fitLL$par[3]))) / (1+exp(fitLL$par[3]))
  DurationMatImmunityDays=exp(fitLL$par[4])
  report_ratio <-  exp(fitLL$par[5])
  hosp3 <- c(report_ratio*hosp2[1:12],report_ratio*hosp2[13]*3)
  
  output <- ode(y=yinit.vector, t=nj_times,  
               func=Novac_dynamics, 
            parms=c(parmset,baseline.txn.rate=baseline.txn.rate,b1=b1,phi=phi,DurationMatImmunityDays=DurationMatImmunityDays))
   t0 <- 108
   al <- nrow(yinit)
   output <- tail(output,t0)
   St <- output[,-1]
   I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
    beta <-  baseline.txn.rate/(parmset$dur.days1/30.44)/(sum(parmset$yinit.matrix)^(1-parmset$q))*parmset$c2# q depends on transmission type (whether depends on population density or not); c2 is the contact matrix
 
    lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}
    
   H1=matrix(0,nrow=t0,ncol=al)#Number of hospitalizations by age
      for (i in 1:al){
     H1[,i]=hosp1[i]*S0[,i]*lambda1[,i]+hosp2[i]*sigma1*S1[,i]*lambda1[,i]+hosp3[i]*sigma2*S2[,i]*lambda1[,i]+hosp3[i]*sigma3*S3[,i]*lambda1[,i]}
  H <- rowSums(H1)
  age_dist <- colSums(H1)
   H <- data.frame(H)
   H$date <- rsv_nj$date
```

## Real Data vs Model Fit
### time series
```{r}
plot(rsv_nj$date,rsv_nj$H,xlab = "date",ylab="RSV cases",type="l")
lines(H$date,H$H,col="red")
```
### Age distribution
```{r}
 # library
library(ggplot2)
library(tidyr)


names(age_dist) <-  agenames

# create a dataset
agedist <- rbind(age_dist/sum(age_dist),nj_agedist/sum(nj_agedist))
agedist <- data.frame(matrix(agedist,ncol=2,nrow=13,byrow = T))
colnames(agedist) <- c("model","data")
agedist$age <- c(1,2,3,4,5,6,7,8,9,10,11,12,13)

data_long <- gather(agedist, group, proportion,model:data, factor_key=TRUE)

 
# Grouped
ggplot(data_long, aes(fill=group, y=proportion, x=age)) + 
    geom_bar(position="dodge", stat="identity")+
  scale_x_discrete(name ="AGE", 
                    limits=agenames)+
  theme_classic()
```

```{r}
rsv_nj$months <- rep(c(7:12,1:6),9)
library(tidyverse)
hosp_cases <- rsv_nj %>%
  group_by(months) %>%
  summarise(rsv=as.integer(mean(rsv)))

stan_initial_y <- as.integer(St[1,])
stan_initial_para <- list(beta=baseline.txn.rate,b1=b1,phi=phi,omega=1/DurationMatImmunityDays,hosp_ratio=report_ratio)
saveRDS(stan_initial_y,"C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/staninput/stan_initial_y.rds")
saveRDS(stan_initial_para,"C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/staninput/stan_initial_para.rds")
data_msis_nj_9yrs <- list(n_months=nrow(rsv_nj),y0=stan_initial_y,t0=0,ts=c(1:nrow(rsv_nj)),N=sum(stan_initial_y),birthrate=c(0.013,rep(0,12)),um=um,rho1=rho1,rho2=rho2,gamma1=gamma1,gamma2=gamma2,gamma3=gamma3,sigma1=sigma1,sigma2=sigma2,sigma3=sigma3,hosp1=hosp1,hosp2=hosp2,c2=contactUSAinfant,q=1,agegroups=13,hosp_age=age_dist,u=1/WidthAgeClassMonth,hosp_cases=rsv_nj$rsv)
saveRDS(data_msis_nj,"C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/staninput/data_msis_nj_9yrs.rds")

data_msis_nj_annual <- list(n_months=12,y0=stan_initial_y,t0=0,ts=1:12,N=sum(stan_initial_y),birthrate=c(0.013,rep(0,12)),um=um,rho1=rho1,rho2=rho2,gamma1=gamma1,gamma2=gamma2,gamma3=gamma3,sigma1=sigma1,sigma2=sigma2,sigma3=sigma3,hosp1=hosp1,hosp2=hosp2,c2=contactUSAinfant,q=1,agegroups=13,hosp_age=age_dist,u=1/WidthAgeClassMonth,hosp_cases=hosp_cases$rsv)
saveRDS(data_msis_nj,"C:/Users/70914/Box/aim3/RSVtransmissionmodel/data_and_parms/staninput/data_msis_nj_annual.rds")
```

